---
title: "Age adjustment vignette"
author: "Shaun Truelove"
date: "5/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root = here::here())
library(tidyverse)
library(covidSeverity)

```

## Load data

```{r}
# # UNHCR data [remove before pushing], add dummy dataset
# age_data <- readr::read_csv("inst/extdata/unhcr_ages.csv")
# age_matrix <- age_data[,c(4:13)]
# rownames(age_matrix) <- age_data$geoid
# #age_data_long <- convert_age_matrix_long(age_matrix)


age_data_long <- bind_rows(
  data.frame(geoid = "010101",
             age_l = c(0, 5, 12, 18, 60, 80),
             age_r = c(5, 12, 18, 60, 80, 100),
             pop = c(112200, 132000, 81000, 252600, 21600, 600), stringsAsFactors = FALSE),
  data.frame(geoid = "020101",
             age_l = c(0, 5, 12, 18, 60, 80),
             age_r = c(5, 12, 18, 60, 80, 100),
             pop = c(110000, 152000, 60000, 202600, 27600, 1200), stringsAsFactors = FALSE))
#age_data <- age_data_long 


```


## Format age data

### For data in a wide or matrix format, with a row of populations by age for each geoid, these can be converted using **convert_age_matrix_long()**:
```{r, eval=FALSE}
age_data_long <- convert_age_matrix_long(age_matrix)
```

### Once age is in long format, we need to convert it to 10-year age groups:
```{r}
age_data10yr <- make_10yr_age_data(age_data_long)
```



## Run age-adjustment
Using the correctly setup data, we can run the age adjustment parameters estimation:
```{r}
geo_age_params <- get_ageadjustments(age_data10yr,
                                     cores=4,
                                     n_sims=40,
                                     n_preds=1000,
                                     age_grps=c(seq(0,80,by=10),100),
                                     googlesheet_access=TRUE,
                                     output_dir="output")
```



## USA OVERALL

```{r, eval=FALSE}

# Get age - already in 10yr age groups
age_data_usa <- get_country_age_pop(ISO3="USA") %>% mutate(geoid="USA")

geo_age_params <- get_ageadjustments( age_data_usa,
                                      cores=4, 
                                      n_sims=40,
                                      n_preds=1000,
                                      age_grps=c(seq(0,80,by=10),100),
                                      googlesheet_access=TRUE,
                                      output_dir="output",
                                      pop_name = "USA")
```

## All Countries
```{r, eval=FALSE}

# Get age - already in 10yr age groups
age_data_world <- get_allcountries_age_pop() %>% mutate(geoid=country)

geo_age_params <- get_ageadjustments(age_data_world,
                                     cores=4, 
                                     n_sims=40,
                                     n_preds=1000,
                                     age_grps=c(seq(0,80,by=10),100),
                                     googlesheet_access=TRUE,
                                     output_dir="output",
                                     pop_name = "AllCountries")
```



# Bangladesh using WorldPop

Users need to download their own shapefiles for this.


### First get the age distributions by Admin level 2

```{r}

filenames <- download_worldpop_agetifs(country="BGD", year="2020", save_dir="../raw_data", cores=4)


age_pop_data <- load_worldpop_age(shp="../raw_data/BGD/bgd_admbnda_adm2_bbs_20180410/bgd_admbnda_adm2_bbs_20180410.shp", 
                                  country="BGD", year="2020", save_dir="../raw_data", cores=4)

age_pop_10yr <- convert_wp_10yr(age_pop_data)
readr::write_csv(age_pop_10yr, "../raw_data/BDG/BDG_age_pop_10yr.csv")
readr::write_csv(age_pop_data, "../raw_data/BDG/BDG_age_pop_data.csv")

```

### Now estimate the severity parameters for all Admin 2 locations in the country
```{r}

# Get proportions
age_pop_10yr <- age_pop_10yr %>% 
    dplyr::group_by(adm2) %>%
    dplyr::mutate(prop = pop / sum(pop)) %>%
    dplyr::ungroup() 

# make a geoid
age_pop_10yr <- age_pop_10yr %>% 
    dplyr::mutate(geoid = as.character(stringr::str_pad(as.integer(as.factor(adm2)), 5, "left", 0)))
readr::write_csv(age_pop_10yr, "../raw_data/BDG/BDG_age_pop_10yr.csv")


# Estimate parameters
geo_age_params <- get_ageadjustments(age_pop_10yr,
                                     cores=4,
                                     n_sims=40,
                                     n_preds=1000,
                                     age_grps=c(seq(0,80,by=10),100),
                                     googlesheet_access=TRUE,
                                     output_dir="../raw_data/BGD",
                                     pop_name=NULL)


```

