---
title: "Age adjustment vignette"
authors: "Stephen Lauer and Shaun Truelove"
date: "5/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root = here::here())
library(tidyverse)
library(covidSeverity)
devtools::install_github("HopkinsIDD/globaltoolboxlite")
library(globaltoolboxlite)

```

## Load data

```{r}
data("US_age_geoid_pct")
data("US_age_geoid_pop")

age_matrix <- US_age_geoid_pop
```


## Format age data

### For data in a wide or matrix format, with a row of populations by age for each geoid, these can be converted using **convert_age_matrix_long()**:
```{r, eval=TRUE}
age_data_long <- convert_age_matrix_long(age_matrix)
```

### Once age is in long format, we need to convert it to 10-year age groups:
 -- These data are already set up correctly in 10-year age groups, so no need to evaluate this step
```{r, eval=FALSE}
#age_data10yr <- make_10yr_age_data(age_data_long)
```
Instead here we just need to rename a column and make proportion
```{r}
age_data10yr <- age_data_long %>% 
  dplyr::rename(age = age_range) %>%
  dplyr::group_by(geoid) %>%
  dplyr::mutate(prop = round(pop / sum(pop),3)) %>%
  dplyr::ungroup()
```



## Run age-adjustment
Using the correctly setup data, we can run the age adjustment parameters estimation:
```{r}
geo_age_params <- get_ageadjustments(age_data10yr,
                                     cores=4,
                                     n_sims=100,
                                     n_preds=1000,
                                     age_grps=c(seq(0,80,by=10),100),
                                     googlesheet_access=TRUE,
                                     output_dir="output")
```



## USA OVERALL

```{r, eval=FALSE}

# Get age - already in 10yr age groups
age_data_usa <- get_country_age_pop(ISO3="USA") %>% mutate(geoid="USA")

geo_age_params <- get_ageadjustments( age_data_usa,
                                      cores=4, 
                                      n_sims=40,
                                      n_preds=1000,
                                      age_grps=c(seq(0,80,by=10),100),
                                      googlesheet_access=TRUE,
                                      output_dir="output",
                                      pop_name = "USA_overall")
```

