---
title: "R Notebook"
output: html_notebook
---

---
title: "Age adjustment vignette"
author: "Shaun Truelove"
date: "5/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root = here::here())
library(tidyverse)
library(covidSeverity)
devtools::install_github("HopkinsIDD/globaltoolboxlite")
library(globaltoolboxlite)

```


# GLOBAL SEVERITY using WorldPop

--DOWNLOAD SHAPEFILES
Users need to download their own shapefiles for this.



### First get the list of countries included in the GADM shapefiles

```{r}

shp_global <- sf::read_sf(shp)

country_codes <- 
country_names <- 


```


### Then get the age distributions by Admin level 2

```{r}

filenames <- globaltoolboxlite::download_worldpop_agetifs(country="BGD", year="2020", save_dir="../raw_data", cores=4)


age_pop_data <- globaltoolboxlite::load_worldpop_age(shp="../raw_data/BGD/bgd_admbnda_adm2_bbs_20180410/bgd_admbnda_adm2_bbs_20180410.shp", 
                                  country="BGD", year="2020", save_dir="../raw_data", cores=4)

age_pop_10yr <- globaltoolboxlite::convert_wp_10yr(age_pop_data)
readr::write_csv(age_pop_10yr, "../raw_data/BGD/BGD_age_pop_10yr.csv")
readr::write_csv(age_pop_data, "../raw_data/BGD/BGD_age_pop_data.csv")

```

### Now estimate the severity parameters for all Admin 2 locations in the country
```{r}

# Get proportions
age_pop_10yr <- age_pop_10yr %>% 
    dplyr::group_by(adm2) %>%
    dplyr::mutate(prop = pop / sum(pop)) %>%
    dplyr::ungroup() 

# make a geoid
age_pop_10yr <- age_pop_10yr %>% 
    dplyr::mutate(geoid = as.character(stringr::str_pad(as.integer(as.factor(adm2)), 5, "left", 0)))
readr::write_csv(age_pop_10yr, "../raw_data/BGD/BGD_age_pop_10yr.csv")


# Estimate parameters
geo_age_params <- get_ageadjustments(age_pop_10yr,
                                     cores=4,
                                     n_sims=40,
                                     n_preds=1000,
                                     age_grps=c(seq(0,80,by=10),100),
                                     googlesheet_access=TRUE,
                                     output_dir="../raw_data/BGD",
                                     pop_name=NULL)


```




### Plot a map of the country, colored by severity metrics

```{r}

library(sf)
library(tidyverse)
library(viridis)

# First generate a geodata.csv like used in the CSP
geodata <- readr::read_csv("../raw_data/BGD/BGD_age_pop_10yr.csv")
geodata <- geodata %>% 
  dplyr::group_by(geoid, adm2) %>%
  dplyr::summarize(pop = round(sum(pop)),
                admin0 = "BGD") %>%
  dplyr::select(geoid, admin2=adm2, admin0, pop)
readr::write_csv(geodata, "../raw_data/BGD/geodata.csv")




map_severity <- function(severity_var = "p_death_inf", 
                         severity_var_label = "IFR",
                         geoid_params = "../raw_data/BGD/geoid-params.csv",
                         geodata = "../raw_data/BGD/geodata.csv",
                         shp = "../raw_data/BGD/bgd_admbnda_adm2_bbs_20180410/bgd_admbnda_adm2_bbs_20180410.shp",
                         shp_name_var = "ADM2_EN",
                         label_locs = NA   ){
  suppressMessages({
    geoid_params_ <- readr::read_csv(geoid_params)
    mapdat <- read_sf(shp) %>%
      dplyr::mutate(NAME = get(shp_name_var))
    geodat <- readr::read_csv(geodata)
  })

  if ("pop" %in% colnames(mapdat)){
    mapdat <- mapdat %>% dplyr::select(-pop)
  }
  if ("geoid" %in% colnames(mapdat)){
    mapdat <- mapdat %>% dplyr::rename(geoid_v0 = geoid)
  }
  
  # Combine data into shapefile
  mapdat <- mapdat %>% 
    dplyr::left_join(geodat, by=c("NAME"="admin2")) %>%
    dplyr::left_join(geoid_params_, by=c("geoid"))
  

  # Plot it
  plt <- ggplot(mapdat) +
            geom_sf(aes(fill = get(severity_var))) +
            theme_void() +
            viridis::scale_fill_viridis() +
            labs(fill=severity_var_label)
  if (!is.na(label_locs)){
    if (tolower(label_locs)=="all"){
      plt <- plt +
        geom_text(mapdat, aes(X,Y, label=NAME), size=5)
     } else if (!is.na(label_locs)){
        label_dat <-  mapdat %>% filter(NAME %in% label_locs)
        plt <- plt + geom_sf_text(data=label_dat, aes(label=NAME), size=3, hjust=0.5)
     }
  }
  
  return(plt)
}



map_severity(severity_var = "p_death_inf", 
             severity_var_label = "IFR",
             geoid_params = "../raw_data/BGD/geoid-params.csv",
             geodata = "../raw_data/BGD/geodata.csv",
             shp = "../raw_data/BGD/bgd_admbnda_adm2_bbs_20180410/bgd_admbnda_adm2_bbs_20180410.shp",
             shp_name_var = "ADM2_EN",
             label_locs = NA)

```

