---
title: "Age adjustment vignette"
author: "Shaun Truelove"
date: "5/1/2020"
output:
  html_document:
    df_print: paged
    fig_caption: true
    theme: lumen
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    pandoc_args: ["--wrap=none", "--top-level-division=chapter"]
  editor_options: 
    chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(root = here::here())
#knitr::opts_chunk$set(dev="CairoPNG")
knitr::opts_chunk$set(dev.args = list(png = list(type = "cairo")))


library(tidyverse)
library(covidSeverity)
#devtools::install_github("HopkinsIDD/globaltoolboxlite", force=TRUE)
library(globaltoolboxlite)
library(foreach)
library(doParallel)
library(gridExtra)
library(sf)


# Functions

filter_var <- function(p_var="p_death_inf", countries_=c("ZMB", "COD","KEN"), data = geoid_params){
  data %>% filter(grepl(paste(countries_, collapse = "|"), geoid)) %>% pull(p_var)
}




```


# GLOBAL SEVERITY using WorldPop

--DOWNLOAD SHAPEFILES
Users need to download their own shapefiles for this.



### First get the list of countries included in the GADM shapefiles

```{r}
getwd()
```

Shapefiles

```{r}

shp <- "raw_data/gadm36_levels_shp/gadm36_2.shp"
shp_adm2 <- sf::read_sf(shp)
"KOS" %in% shp_adm2$GID_0

shp1 <- "raw_data/gadm36_levels_shp/gadm36_1.shp"
shp_adm1 <- sf::read_sf(shp1)
"KOS" %in% shp_adm1$GID_0
"XKO" %in% shp_adm1$GID_0

shp0 <- "raw_data/gadm36_levels_shp/gadm36_0.shp"
shp_adm0 <- sf::read_sf(shp0)
"KOS" %in% shp_adm0$GID_0


```



```{r}

shp_adm2 <- shp_adm2 %>% mutate(GID_0 = ifelse(GID_0=="XKO", "KOS", GID_0)) # Fix Kosovo
st_write(shp_adm2, shp, delete_layer=TRUE, append = FALSE) # Re-save it

shp_adm0 <- shp_adm0 %>% mutate(GID_0 = ifelse(GID_0=="XKO", "KOS", GID_0)) # Fix Kosovo
st_write(shp_adm0, shp0, delete_layer=TRUE, append = FALSE) # Re-save it

# For countries with no Level 2 shapefiles, grab the level 1 or 0
shp_adm1 <- shp_adm1 %>% mutate(GID_0 = ifelse(GID_0=="XKO", "KOS", GID_0)) # Fix Kosovo
st_write(shp_adm1, shp1, delete_layer=TRUE, append = FALSE) # Re-save it


shp_adm2_countries <- unique(shp_adm2$GID_0)
shp_adm0_countries <- unique(shp_adm0$GID_0)
shp_adm1_countries <- unique(shp_adm1$GID_0)


adm1only_countries <- shp_adm1_countries[!(shp_adm1_countries %in% shp_adm2_countries)]
adm0only_countries <- shp_adm0_countries[!(shp_adm0_countries %in% c(adm1only_countries, shp_adm2_countries))]


shp_global_adm2 <- shp_adm2  %>% mutate(source = "gadm36_2") %>%
  bind_rows(shp_adm1 %>% filter(GID_0 %in% adm1only_countries) %>% mutate(source = "gadm36_1")) %>%
                                  bind_rows(shp_adm0 %>% filter(GID_0 %in% adm0only_countries) %>% mutate(source = "gadm36_0"))            

# for simplicity, i will rename the 
shp_global_adm2 <- shp_global_adm2 %>% 
  mutate(NAME_2 = ifelse(is.na(GID_2) & !is.na(GID_1), NAME_1, NAME_2),
         GID_2 = ifelse(is.na(GID_2) & !is.na(GID_1), GID_1, GID_2)) %>%
  mutate(NAME_2 = ifelse(is.na(GID_2) & !is.na(GID_0), NAME_0, NAME_2),
         GID_2 = ifelse(is.na(GID_2) & !is.na(GID_0), GID_0, GID_2))

# standardize the names of adm2
shp_global_adm2 <- shp_global_adm2 %>%
  mutate(NAME_2_orig = NAME_2,
         NAME_1_orig = NAME_1,
         NAME_2 = globaltoolboxlite::standardize_location_strings(NAME_2),
         NAME_1 = globaltoolboxlite::standardize_location_strings(NAME_1))

#shp_global_adm2_afg <- sf::read_sf(shp) %>% filter(!!as.name(shp_country_var)=="AFG")
shp_data <- sf::st_set_geometry(shp_global_adm2, NULL)

country_codes_gadm <- unique(shp_global_adm2$GID_0)
country_names <- unique(shp_global_adm2$NAME_0)
country_codes <- country_codes_gadm
#country_codes[country_codes=="XKO"] <- "KOS"

# Save the combined shapefiles
st_write(shp_global_adm2, "raw_data/gadm36_levels_shp/gadm36_combo.shp", delete_layer=TRUE, append = FALSE) # Re-save it


```


### Then get the age distributions by Admin level 2

```{r, eval=FALSE}

dir.create(file.path("raw_data", "global_severity"), recursive = TRUE, showWarnings = FALSE)

for (c in seq_along(country_codes)){ 
#for (c in 163:length(country_codes)){ 
  
  # Download and save all the geotifs from worldpop
  filenames <- globaltoolboxlite::download_worldpop_agetifs(country=country_codes[c], year="2020", save_dir="raw_data/geotifs", cores=25)
  
  age_pop_data <- globaltoolboxlite::load_worldpop_age(shp=shp_global_adm2, 
                                                       country=country_codes[c], year="2020", 
                                                       save_dir="raw_data/geotifs", 
                                                       cores=25,
                                                       loc_var = "NAME_2",
                                                       add_pop_to_shapefile = FALSE,
                                                       shp_country_var="GID_0")
  
  age_pop_data <- age_pop_data %>% dplyr::rename(adm2=NAME_2)
  age_pop_10yr <- globaltoolboxlite::convert_wp_10yr(age_pop_data)
  readr::write_csv(age_pop_10yr, file.path("raw_data", "global_severity", paste0(country_codes[c],"_age_pop_10yr.csv")))
  readr::write_csv(age_pop_data, file.path("raw_data", "global_severity", paste0(country_codes[c],"_age_pop_data.csv")))
  
  # delete the geotifs to save space (they are >10G per country)
  unlink(file.path("../raw_data/geotifs", country_codes[c]), recursive = TRUE)
}

```



### Get distributions for KOS and countries without ADm2 shapefiles

```{r, eval=FALSE}


# Kosovo
# - renamed shapefile GID_0 above

country_codes_tmp <- "KOS"
{
 # Download and save all the geotifs from worldpop
  filenames <- globaltoolboxlite::download_worldpop_agetifs(country=country_codes_tmp, year="2020", save_dir="raw_data/geotifs", cores=25)

  age_pop_data <- globaltoolboxlite::load_worldpop_age(shp=shp_global_adm2, 
                                                       country=country_codes_tmp, year="2020", 
                                                       save_dir="raw_data/geotifs", 
                                                       cores=25,
                                                       loc_var = "NAME_2",
                                                       add_pop_to_shapefile = FALSE,
                                                       shp_country_var="GID_0")
  
  age_pop_data <- age_pop_data %>% dplyr::rename(adm2=NAME_2)
  age_pop_10yr <- globaltoolboxlite::convert_wp_10yr(age_pop_data)
  readr::write_csv(age_pop_10yr, file.path("raw_data", "global_severity", paste0(country_codes_tmp,"_age_pop_10yr.csv")))
  readr::write_csv(age_pop_data, file.path("raw_data", "global_severity", paste0(country_codes_tmp,"_age_pop_data.csv")))
  
  # delete the geotifs to save space (they are >10G per country)
  unlink(file.path("raw_data/geotifs", country_codes_tmp), recursive = TRUE)
}




# Get countries without Adm2

# Adm1 only countries
for (c in 4:length(adm1only_countries)){ 
  
  # Download and save all the geotifs from worldpop
  filenames <- tryCatch({
    globaltoolboxlite::download_worldpop_agetifs(country=adm1only_countries[c], 
                                                 year="2020", save_dir="raw_data/geotifs", cores=25)
  },
  error=function(e) NA
  )
  if (is.na(filenames)) next
  
  age_pop_data <- tryCatch({
    globaltoolboxlite::load_worldpop_age(shp=shp_global_adm2, 
                                                       country=adm1only_countries[c], 
                                                       year="2020", 
                                                       save_dir="raw_data/geotifs", 
                                                       cores=25,
                                                       loc_var = "NAME_2",
                                                       add_pop_to_shapefile = FALSE,
                                                       shp_country_var="GID_0")
  },
  error=function(e) NA
  )
  if (is.na(age_pop_data)) next
  
  age_pop_data <- age_pop_data %>% dplyr::rename(adm2=NAME_2)
  age_pop_10yr <- globaltoolboxlite::convert_wp_10yr(age_pop_data)
  readr::write_csv(age_pop_10yr, file.path("raw_data", "global_severity", paste0(adm1only_countries[c],"_age_pop_10yr.csv")))
  readr::write_csv(age_pop_data, file.path("raw_data", "global_severity", paste0(adm1only_countries[c],"_age_pop_data.csv")))
  
  # delete the geotifs to save space (they are >10G per country)
  unlink(file.path("raw_data/geotifs", adm1only_countries[c]), recursive = TRUE)
}


# Adm0 only countries
for (c in 1:length(adm0only_countries)){ 
  
  # Download and save all the geotifs from worldpop
  filenames <- tryCatch({
    globaltoolboxlite::download_worldpop_agetifs(country=adm0only_countries[c], 
                                                            year="2020", save_dir="raw_data/geotifs", cores=25)
  },
    error=function(e) NA
  )
  if (is.na(filenames)) next
  
  age_pop_data <- tryCatch({
    globaltoolboxlite::load_worldpop_age(shp=shp_global_adm2, 
                                                       country=adm0only_countries[c], year="2020", 
                                                       save_dir="raw_data/geotifs", 
                                                       cores=25,
                                                       loc_var = "NAME_2",
                                                       add_pop_to_shapefile = FALSE,
                                                       shp_country_var="GID_0")
  },
  error=function(e) NA
  )
  if (is.na(age_pop_data)) next
  
  
  age_pop_data <- age_pop_data %>% dplyr::rename(adm2=NAME_2)
  age_pop_10yr <- globaltoolboxlite::convert_wp_10yr(age_pop_data)
  readr::write_csv(age_pop_10yr, file.path("raw_data", "global_severity", paste0(adm0only_countries[c],"_age_pop_10yr.csv")))
  readr::write_csv(age_pop_data, file.path("raw_data", "global_severity", paste0(adm0only_countries[c],"_age_pop_data.csv")))
  
  # delete the geotifs to save space (they are >10G per country)
  unlink(file.path("raw_data/geotifs", adm0only_countries[c]), recursive = TRUE)
}



```






### Now estimate the severity parameters for all Admin 2 locations in each country

```{r, eval=FALSE}

#for (c in seq_along(country_codes)){ 
for (c in 170:length(country_codes)){ 

  if (!file.exists(file.path("raw_data", "global_severity", paste0(country_codes[c],"_age_pop_10yr.csv")))) next
  #if (country_codes[c]=="KOS" | country_codes[c]=="XKO") next
  age_pop_10yr <- readr::read_csv(file.path("raw_data", "global_severity", paste0(country_codes[c],"_age_pop_10yr.csv")))
  
  # Get rid of .y columns and fix .x
  age_pop_10yr <- age_pop_10yr %>%
     dplyr::select(-contains(".y"))
  colnames(age_pop_10yr) <- gsub(".x", "", colnames(age_pop_10yr))
   
  # Get proportions
  age_pop_10yr <- age_pop_10yr %>% 
    dplyr::select("adm2","pop","pop_m", "pop_f", "age_l","age_r", "age") %>%
    dplyr::mutate(adm2 = globaltoolboxlite::standardize_location_strings(adm2)) %>%
    dplyr::group_by(adm2) %>%
    dplyr::mutate(prop = round(pop / sum(pop),4)) %>%
    dplyr::ungroup() 
  
  # make a geoid and province info
  age_pop_10yr <- age_pop_10yr %>% 
    #dplyr::mutate(geoid = as.character(iconvadm2)), 5, "left", 0)))
    left_join(shp_data %>% 
                dplyr::filter(GID_0==country_codes[c]) %>%
                dplyr::select(adm0=GID_0, country=NAME_0, 
                              adm1=GID_1, adm1_name=NAME_1, 
                              geoid=GID_2, adm2=NAME_2), 
              by=c("adm2"="adm2"))
  
  # Fix 0s in proportion when 0 population in a district
  age_pop_10yr <- age_pop_10yr %>% 
    dplyr::mutate(prop = ifelse(is.nan(prop), 0, prop))
  
  # Save the updated version
  readr::write_csv(age_pop_10yr, file.path("raw_data", "global_severity", paste0(country_codes[c],"_age_pop_10yr.csv")))
  
  if (file.exists(file.path("raw_data", "global_severity", paste0(country_codes[c],"_geoid-params.csv")))) next
  
  # Estimate parameters
  geo_age_params <- get_ageadjustments(age_pop_10yr,
                                       cores=25,
                                       n_sims=40,
                                       n_preds=1000,
                                       age_grps=c(seq(0,80,by=10),100),
                                       googlesheet_access=FALSE,
                                       output_dir="raw_data/global_severity",
                                       pop_name=country_codes[c])
}

```




### Plot a map of the countries, colored by severity metrics

```{r}
source("R/plotting_functions.R")

```



```{r, eval=FALSE}

# Merge results

geoid_params <- age_pop_10yr <- list()

for (c in seq_along(country_codes)){ 
  #if (country_codes[c]=="KOS" | country_codes[c]=="XKO") next
  suppressMessages({
    geoid_params[[c]] <- tryCatch( readr::read_csv(file.path("raw_data", "global_severity", paste0(country_codes[c],"_geoid-params.csv"))), error=function(e) NULL)
    age_pop_10yr[[c]] <- tryCatch( readr::read_csv(file.path("raw_data", "global_severity", paste0(country_codes[c],"_age_pop_10yr.csv"))), error=function(e) NULL)
  })
}

geoid_params <- data.table::rbindlist(geoid_params)
age_pop_10yr <- data.table::rbindlist(age_pop_10yr)


# generate a geodata.csv like used in the CSP
geodata <- age_pop_10yr %>% 
  dplyr::group_by(geoid, adm2, adm0, adm1, country, adm1_name) %>%
  dplyr::summarize(pop = round(sum(pop)),
                   admin0 = country_codes[c]) %>%
  dplyr::select(geoid, adm2=adm2, adm1=adm1, adm0=adm0, adm1_name, country, pop)

# Add regions to geodata
geodata <- geodata %>%
  dplyr::mutate(region = globaltoolboxlite::get_region(adm0),
                who_region = globaltoolboxlite::get_whoregion(adm0),
                subregion = globaltoolboxlite::get_subregion(adm0))

geodata <- geodata %>% dplyr::rename(admin2=adm2, admin1=adm1, admin0=adm0) # so they match with the CSP


readr::write_csv(geoid_params, file.path("raw_data", "global_severity", "GLOBAL_geoid_params.csv"))
readr::write_csv(age_pop_10yr, file.path("raw_data", "global_severity", "GLOBAL_age_pop_10yr.csv"))
readr::write_csv(geodata, file.path("raw_data", "global_severity", "GLOBAL_geodata.csv"))

```



# ---> need to figure out what happens to RR with pop=0 <-----



```{r}

geoid_params <- readr::read_csv(file.path("raw_data", "global_severity", "GLOBAL_geoid_params.csv"))
age_pop_10yr <- readr::read_csv(file.path("raw_data", "global_severity", "GLOBAL_age_pop_10yr.csv"))
geodata <- readr::read_csv(file.path("raw_data", "global_severity", "GLOBAL_geodata.csv"))

table(geodata$region)

# Make locations with 0 population estimates NA
geoid_params <- geoid_params %>% 
  dplyr::left_join(geodata %>% dplyr::select(geoid, pop))

cols_to_na <- which(!(colnames(geoid_params) %in% c("geoid","pop")))
which_na <- which(geoid_params$pop==0)
geoid_params[which_na, cols_to_na] <- NA


# add some vars to geoid_params
geoid_params <- geoid_params %>% 
  dplyr::mutate(p_asymp_inf = 1 - p_symp_inf,
                p_asymp_mild_inf = p_asymp_inf + p_mild_inf)


mapdat <- merge_severity_gadm_data(
                              geoid_params = geoid_params,
                              geodata = geodata,
                              shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_combo.shp",
                              shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                              shp2_name_var = "GID_2")



```







# Build the map
We will start small because it takes a while to load



# Proportion not asympt or mild (essentially detectible cases)


```{r}

mapdat$shp_adm2$p_not_asymptmild_inf <- 1 - mapdat$shp_adm2$p_asymp_mild_inf
mapdat_data <- sf::st_set_geometry(mapdat$shp_adm2, NULL)

hist(mapdat$shp_adm2$p_not_asymptmild_inf, breaks=50)
hist(filter_var("p_not_asymptmild_inf", countries_, mapdat$shp_adm2), breaks=50)
hist(filter_var("p_not_asymptmild_inf", "USA", mapdat$shp_adm2), breaks=50)


p_not_asympt_mild <- map_severity_gadm_fast(severity_var = "p_not_asymptmild_inf", 
                  severity_var_label = "Pr(Not Asympt or Mild)",
                  .region = "Africa",
                  .country = NULL,
                  merged_data = mapdat,
                  color_pal = "inferno",
                  legend_pos = c(.2, .2),
                  leg_breaks = seq(0,.2,.05),
                  log_trans = FALSE,
                  leg_range = c(0,.2),
                  border_size = c(.25, .01),
                  label_locs = NA) +
                  coord_sf(xlim = c(-20, 53), ylim=c(-35,40), expand=FALSE) +
                  theme(legend.text=element_text(size=rel(0.5))) +
                  guides(colour=guide_legend(override.aes = list(size = .75)))


#plot(p_not_asympt_mild)
#ggsave("raw_data/figures/p_asymptmild_africa.pdf", plot=p_not_asympt_mild, dpi=600)
ggsave("raw_data/figures/p_notasymptmild_africa.png", plot=p_not_asympt_mild, dpi=600)


p_not_asympt_mild_b <- map_severity_gadm_fast(severity_var = "p_not_asymptmild_inf", 
                  severity_var_label = "Pr(Not Asympt or Mild)",
                  .region = "Americas",
                  region_var = "region",
                  .country = c("USA","CAN","MEX", globaltoolboxlite::get_iso(c("Guatemala","Nicaragua","Honduras","Costa Rica","El Salvador","Belize","Panama"))),
                  color_pal = "inferno",
                  merged_data = mapdat,
                  legend_pos = c(.2,.2),
                  leg_breaks = seq(0,.2,.05),
                  log_trans = FALSE,
                  leg_range = c(0,.2),
                  border_size = c(.25, .01),
                  label_locs = NA) +
                  coord_sf(xlim = c(-180, -55), ylim=c(10,75), expand=FALSE) +
                  theme(legend.text=element_text(size=rel(0.5))) +
                  guides(colour=guide_legend(override.aes = list(size = .75)))

#plot(p_not_asympt_mild_b)
#ggsave("raw_data/figures/p_asymptmild_usa.pdf", plot=p_not_asympt_mild_b, dpi=300)
ggsave("raw_data/figures/p_notasymptmild_NAmerica.png", plot=p_not_asympt_mild_b, dpi=600)


p_comb <- gridExtra::grid.arrange(p_not_asympt_mild_b +        
                                    theme(legend.position = c(.4,.2)) +
                                    guides(colour=guide_legend(override.aes = list(size = .05))),
                                  p_not_asympt_mild +
                                    theme(legend.position = "none"), 
                                  ncol=2, nrow=1, widths=c(1.3,1))

ggsave("raw_data/figures/p_notasymptmild_NAmerica_Africa.png", plot=p_comb, dpi=600, width = 14, height = 8, units = "in")

```























```{r}


countries_ <- get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC"))

par(mfrow=c(2,1))
hist(geoid_params$p_death_inf, breaks=50)
hist(filter_var("p_death_inf", countries_), breaks=50)

hist(geoid_params$p_symp_inf, breaks=50)
hist(filter_var("p_symp_inf", countries_), breaks=50)
summary(filter_var("p_symp_inf", countries_))

hist(geoid_params$p_asymp_mild_inf, breaks=50)
hist(filter_var("p_asymp_mild_inf", countries_), breaks=50)
par(mfrow=c(1,1))


p_ifr <- map_severity_gadm(severity_var = "p_death_inf", 
                  severity_var_label = "IFR",
                  .region = "Africa",
                  .country = countries_,
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = shp_adm2,
                  shp_adm0 = shp_adm0,
                  shp2_name_var = "GID_2",
                  legend_pos = c(.9,.1),
                  leg_breaks = c(0, 0.001, 0.002, 0.003, 0.004, 0.005, 0.007, 0.01),
                  log_trans = TRUE,
                  leg_range = c(0,.01),
                  label_locs = NA)
plot(p_ifr)
#ggsave("raw_data/figures/p_ifr_africa.pdf", plot=p_ifr, dpi=600)
ggsave("raw_data/figures/p_ifr_africa.png", plot=p_ifr, dpi=300)

# Faster version using pre-merged data
p_ifr <- map_severity_gadm_fast(severity_var = "p_death_inf", 
                  severity_var_label = "IFR",
                  .region = "Africa",
                  .country = countries_,
                  merged_data = mapdat,
                  legend_pos = c(.9,.1),
                  leg_breaks = c(0, 0.001, 0.002, 0.003, 0.004, 0.005, 0.007, 0.01),
                  log_trans = TRUE,
                  leg_range = c(0,.01),
                  label_locs = NA)

plot(p_ifr)
#ggsave("raw_data/figures/p_ifr_africa.pdf", plot=p_ifr, dpi=600)
ggsave("raw_data/figures/p_ifr_africa.png", plot=p_ifr, dpi=300)




p_sympt <- map_severity_gadm_fast(severity_var = "p_symp_inf", 
                  severity_var_label = "Pr(Sympt)",
                  .region = "Africa",
                  .country = countries_,
                  merged_data = mapdat,
                  legend_pos = c(.8,.8),
                  leg_breaks = seq(.3,.5, .05),
                  log_trans = FALSE,
                  leg_range = c(.345,.45),
                  label_locs = NA)
plot(p_sympt)
#ggsave("raw_data/figures/p_sympt_africa.pdf", plot=p_sympt, dpi=600)
ggsave("raw_data/figures/p_sympt_africa.png", plot=p_sympt, dpi=300)






p_asympt_mild <- map_severity_gadm(severity_var = "p_asymp_mild_inf", 
                  severity_var_label = "Pr(Asympt or Mild)",
                  .region = "Africa",
                  .country = countries_,
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = shp_adm2,
                  shp_adm0 = shp_adm0,
                  shp2_name_var = "GID_2",
                  legend_pos = c(.9,.15),
                  leg_breaks = seq(.80,1,.05),
                  log_trans = FALSE,
                  leg_range = c(.80,1),
                  label_locs = NA)

plot(p_asympt_mild)
ggsave("raw_data/figures/p_asymptmild_africa.pdf", plot=p_asympt_mild, dpi=600)
ggsave("raw_data/figures/p_asymptmild_africa.png", plot=p_asympt_mild, dpi=600)



p_asympt_mild <- map_severity_gadm(severity_var = "p_asymp_mild_inf", 
                  severity_var_label = "Pr(Asympt or Mild)",
                  .region = "Americas",
                  .country = "USA",
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = shp_adm2,
                  shp_adm0 = shp_adm0,
                  shp2_name_var = "GID_2",
                  legend_pos = c(.9,.15),
                  leg_breaks = seq(.80,1,.05),
                  log_trans = FALSE,
                  leg_range = c(.80,1),
                  label_locs = NA) +
                  coord_sf(xlim = c(-150, -60), expand=FALSE)

plot(p_asympt_mild)
ggsave("raw_data/figures/p_asymptmild_usa.pdf", plot=p_asympt_mild, dpi=600)
ggsave("raw_data/figures/p_asymptmild_usa.png", plot=p_asympt_mild, dpi=600)

summary(filter_var("p_asymp_mild_inf", countries_))
summary(filter_var("p_asymp_mild_inf", "USA"))


```



```{r}


summary(filter_var("p_not_asymptmild_inf", countries_, mapdat$shp_adm2))
summary(filter_var("p_not_asymptmild_inf", "USA", mapdat$shp_adm2))













# Americas
summary(geoid_params$p_asymp_mild_inf)
hist(geoid_params$p_asymp_mild_inf, breaks=100)
#tmp <- as.data.frame(table(geoid_params$p_asymp_mild_inf))

p_asympt_mild <- map_severity_gadm(severity_var = "p_asymp_mild_inf", 
                  severity_var_label = "Pr(Asympt or Mild)",
                  .region = "Americas",
                  .country = NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, .05,.10,.2,.3,.35,.4,.45,.5,.55,.6,.8,1),
                  log_trans = FALSE,
                  leg_range = c(.7,1),
                  label_locs = NA)

ggsave("raw_data/figures/p_asymptmild_americas.pdf", plot=p_asympt_mild, dpi=600)
ggsave("raw_data/figures/p_asymptmild_americas.png", plot=p_asympt_mild, dpi=600)




# Europe
summary(geoid_params$p_asymp_mild_inf)
hist(geoid_params$p_asymp_mild_inf, breaks=100)
#tmp <- as.data.frame(table(geoid_params$p_asymp_mild_inf))

p_asympt_mild <- map_severity_gadm(severity_var = "p_asymp_mild_inf", 
                  severity_var_label = "Pr(Asympt or Mild)",
                  .region = "Europe",
                  .country = NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, .05,.10,.2,.3,.35,.4,.45,.5,.55,.6,.8,1),
                  log_trans = FALSE,
                  leg_range = c(.7,1),
                  label_locs = NA)

ggsave("raw_data/figures/p_asymptmild_europe.pdf", plot=p_asympt_mild, dpi=600)
ggsave("raw_data/figures/p_asymptmild_europe.png", plot=p_asympt_mild, dpi=600)






```








```{r}

hist(geoid_params$p_death_inf, breaks=50)

p_ifr <- map_severity_gadm(severity_var = "p_death_inf", 
                  severity_var_label = "IFR",
                  .region = "Africa",
                  .country = NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, 0.001, 0.002, 0.003, 0.004, 0.005, 0.007, 0.01),
                  log_trans = TRUE,
                  leg_range = c(0,.01),
                  label_locs = NA)

dir.create("raw_data/figures",recursive = TRUE)
ggsave("raw_data/figures/p_ifr_africa.pdf", plot=p_ifr, dpi=600)
ggsave("raw_data/figures/p_ifr_africa.png", plot=p_ifr, dpi=600)



summary(geoid_params$p_symp_inf)
hist(geoid_params$p_symp_inf, breaks=100)

p_sympt <- map_severity_gadm(severity_var = "p_symp_inf", 
                  severity_var_label = "Pr(Sympt)",
                  .region = "Africa",
                  .country = NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, .05,.10,.2,.3,.35,.4,.45,.5,.55,.6,.8,1),
                  log_trans = FALSE,
                  leg_range = c(.3,.75),
                  label_locs = NA)

ggsave("raw_data/figures/p_sympt_africa.pdf", plot=p_sympt, dpi=600)
ggsave("raw_data/figures/p_sympt_africa.png", plot=p_sympt, dpi=600)




summary(geoid_params$p_asymp_mild_inf)
hist(geoid_params$p_asymp_mild_inf, breaks=100)
#tmp <- as.data.frame(table(geoid_params$p_asymp_mild_inf))

p_asympt_mild <- map_severity_gadm(severity_var = "p_asymp_mild_inf", 
                  severity_var_label = "Pr(Asympt or Mild)",
                  .region = "Africa",
                  .country = NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, .05,.10,.2,.3,.35,.4,.45,.5,.55,.6,.8,1),
                  log_trans = FALSE,
                  leg_range = c(.7,1),
                  label_locs = NA)

ggsave("raw_data/figures/p_asymptmild_africa.pdf", plot=p_asympt_mild, dpi=600)
ggsave("raw_data/figures/p_asymptmild_africa.png", plot=p_asympt_mild, dpi=600)







# Americas
summary(geoid_params$p_asymp_mild_inf)
hist(geoid_params$p_asymp_mild_inf, breaks=100)
#tmp <- as.data.frame(table(geoid_params$p_asymp_mild_inf))

p_asympt_mild <- map_severity_gadm(severity_var = "p_asymp_mild_inf", 
                  severity_var_label = "Pr(Asympt or Mild)",
                  .region = "Americas",
                  .country = NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, .05,.10,.2,.3,.35,.4,.45,.5,.55,.6,.8,1),
                  log_trans = FALSE,
                  leg_range = c(.7,1),
                  label_locs = NA)

ggsave("raw_data/figures/p_asymptmild_americas.pdf", plot=p_asympt_mild, dpi=600)
ggsave("raw_data/figures/p_asymptmild_americas.png", plot=p_asympt_mild, dpi=600)




# Europe
summary(geoid_params$p_asymp_mild_inf)
hist(geoid_params$p_asymp_mild_inf, breaks=100)
#tmp <- as.data.frame(table(geoid_params$p_asymp_mild_inf))

p_asympt_mild <- map_severity_gadm(severity_var = "p_asymp_mild_inf", 
                  severity_var_label = "Pr(Asympt or Mild)",
                  .region = "Europe",
                  .country = NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, .05,.10,.2,.3,.35,.4,.45,.5,.55,.6,.8,1),
                  log_trans = FALSE,
                  leg_range = c(.7,1),
                  label_locs = NA)

ggsave("raw_data/figures/p_asymptmild_europe.pdf", plot=p_asympt_mild, dpi=600)
ggsave("raw_data/figures/p_asymptmild_europe.png", plot=p_asympt_mild, dpi=600)






```





# Sub Regions    

```{r}

geoid_params <- readr::read_csv(file.path("../raw_data", "global_severity", "GLOBAL_geoid_params.csv"))
age_pop_10yr <- readr::read_csv(file.path("../raw_data", "global_severity", "GLOBAL_age_pop_10yr.csv"))
geodata <- readr::read_csv(file.path("../raw_data", "global_severity", "GLOBAL_geodata.csv"))

# Make locations with 0 population estimates NA
geoid_params <- geoid_params %>% 
  dplyr::left_join(geodata %>% dplyr::select(geoid, pop, region, who_region, subregion))

cols_to_na <- which(!(colnames(geoid_params) %in% c("geoid","pop")))
which_na <- which(geoid_params$pop==0)
geoid_params[which_na, cols_to_na] <- NA

# add some vars to geoid_params
geoid_params <- geoid_params %>% 
  dplyr::mutate(p_asymp_inf = 1 - p_symp_inf,
                p_asymp_mild_inf = p_asymp_inf + p_mild_inf)
geoid_params_plusregions <- geoid_params
geoid_params <- geoid_params %>% dplyr::select(-c(region, subregion, who_region, pop))


table(geodata$region)
table(geodata$subregion)
table(geodata$who_region)
```



```{r}
# Build the map
hist(geoid_params$p_death_inf, breaks=50)

reg_ <- "Northern America"
reg_var <- "subregion"

hist((geoid_params %>% filter(get(reg_var) %in% reg_))$p_death_inf, breaks=50)


p_ifr <- map_severity_gadm(severity_var = "p_death_inf", 
                  severity_var_label = "IFR",
                  .region = reg_,
                  region_var = reg_var,
                  .country = NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, 0.001, 0.002, 0.003, 0.004, 0.005, 0.007, 0.01, 0.02, 0.03),
                  log_trans = FALSE,
                  color_pal = "plasma",
                  label_locs = NA)
p_ifr
dir.create("raw_data/figures",recursive = TRUE, showWarnings = FALSE)
ggsave(paste0("raw_data/figures/p_ifr_", reg_, ".pdf"), plot=p_ifr, dpi=600)
ggsave(paste0("raw_data/figures/p_ifr_", reg_, ".png"), plot=p_ifr, dpi=600)

```



```{r}
reg_ <- " Sub-Saharan Africa"
reg_var <- "subregion"

p_ifr <- map_severity_gadm(severity_var = "p_death_inf", 
                  severity_var_label = "IFR",
                  .region = reg_,
                  region_var = reg_var,
                  .country = NULL, #NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, 0.001, 0.002, 0.003, 0.004, 0.005, 0.007, 0.01, 0.02, 0.03),
                  log_trans = FALSE,
                  leg_range = c(0,.015),
                  color_pal = "plasma",
                  label_locs = NA)

dir.create("raw_data/figures",recursive = TRUE, showWarnings = FALSE)
ggsave(paste0("raw_data/figures/p_ifr_", reg_, ".pdf"), plot=p_ifr, dpi=600)
ggsave(paste0("raw_data/figures/p_ifr_", reg_, ".png"), plot=p_ifr, dpi=600)




```


# Run global

```{r}


geoid_params <- readr::read_csv(file.path("../raw_data", "global_severity", "GLOBAL_geoid_params.csv"))
age_pop_10yr <- readr::read_csv(file.path("../raw_data", "global_severity", "GLOBAL_age_pop_10yr.csv"))
geodata <- readr::read_csv(file.path("../raw_data", "global_severity", "GLOBAL_geodata.csv"))

# Make locations with 0 population estimates NA
geoid_params <- geoid_params %>% 
  dplyr::left_join(geodat %>% dplyr::select(geoid, pop))

cols_to_na <- which(!(colnames(geoid_params) %in% c("geoid","pop")))
which_na <- which(geoid_params$pop==0)
geoid_params[which_na, cols_to_na] <- NA

# add some vars to geoid_params
geoid_params <- geoid_params %>% 
  dplyr::mutate(p_asymp_inf = 1 - p_symp_inf,
                p_asymp_mild_inf = p_asymp_inf + p_mild_inf)


table(geodata$region)
table(geodata$subregion)
table(geodata$who_region)





# Build the map
hist(geoid_params$p_death_inf, breaks=50)

p_ifr <- map_severity_gadm(severity_var = "p_death_inf", 
                  severity_var_label = "IFR",
                  .region = "Africa",
                  region_var = "region",
                  .country = "ZMB", #NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, 0.001, 0.002, 0.003, 0.004, 0.005, 0.007, 0.01, 0.02, 0.03),
                  log_trans = TRUE,
                  color_pal = "plasma",
                  label_locs = NA)

dir.create("raw_data/figures",recursive = TRUE)
ggsave("raw_data/figures/p_ifr_africa.pdf", plot=p_ifr, dpi=600)
ggsave("raw_data/figures/p_ifr_africa.png", plot=p_ifr, dpi=600)



summary(geoid_params$p_symp_inf)
hist(geoid_params$p_symp_inf, breaks=100)

p_sympt <- map_severity_gadm(severity_var = "p_symp_inf", 
                  severity_var_label = "Pr(Sympt)",
                  .region = "Africa",
                  .country = NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, .05,.10,.2,.3,.35,.4,.45,.5,.55,.6,.8,1),
                  log_trans = FALSE,
                  leg_range = c(.3,.75),
                  label_locs = NA)

ggsave("raw_data/figures/p_sympt_africa.pdf", plot=p_sympt, dpi=600)
ggsave("raw_data/figures/p_sympt_africa.png", plot=p_sympt, dpi=600)




summary(geoid_params$p_asymp_mild_inf)
hist(geoid_params$p_asymp_mild_inf, breaks=100)
#tmp <- as.data.frame(table(geoid_params$p_asymp_mild_inf))

p_asympt_mild <- map_severity_gadm(severity_var = "p_asymp_mild_inf", 
                  severity_var_label = "Pr(Asympt or Mild)",
                  .region = "Africa",
                  .country = NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, .05,.10,.2,.3,.35,.4,.45,.5,.55,.6,.8,1),
                  log_trans = FALSE,
                  leg_range = c(.7,1),
                  label_locs = NA)

ggsave("raw_data/figures/p_asymptmild_africa.pdf", plot=p_asympt_mild, dpi=600)
ggsave("raw_data/figures/p_asymptmild_africa.png", plot=p_asympt_mild, dpi=600)







# Americas
summary(geoid_params$p_asymp_mild_inf)
hist(geoid_params$p_asymp_mild_inf, breaks=100)
#tmp <- as.data.frame(table(geoid_params$p_asymp_mild_inf))

p_asympt_mild <- map_severity_gadm(severity_var = "p_asymp_mild_inf", 
                  severity_var_label = "Pr(Asympt or Mild)",
                  .region = "Americas",
                  .country = NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, .05,.10,.2,.3,.35,.4,.45,.5,.55,.6,.8,1),
                  log_trans = FALSE,
                  leg_range = c(.7,1),
                  label_locs = NA)

ggsave("raw_data/figures/p_asymptmild_americas.pdf", plot=p_asympt_mild, dpi=600)
ggsave("raw_data/figures/p_asymptmild_americas.png", plot=p_asympt_mild, dpi=600)




# Europe
summary(geoid_params$p_asymp_mild_inf)
hist(geoid_params$p_asymp_mild_inf, breaks=100)
#tmp <- as.data.frame(table(geoid_params$p_asymp_mild_inf))

p_asympt_mild <- map_severity_gadm(severity_var = "p_asymp_mild_inf", 
                  severity_var_label = "Pr(Asympt or Mild)",
                  .region = "Europe",
                  .country = NULL, #get_iso(c("Zambia","Tanzania", "Uganda","Rwanda","Kenya", "DRC")),
                  geoid_params = geoid_params,
                  geodata = geodata,
                  shp_adm2 = "raw_data/gadm36_levels_shp/gadm36_2.shp",
                  shp_adm0 = "raw_data/gadm36_levels_shp/gadm36_0.shp",
                  shp2_name_var = "GID_2",
                  legend_pos = c(.8,.8),
                  leg_breaks = c(0, .05,.10,.2,.3,.35,.4,.45,.5,.55,.6,.8,1),
                  log_trans = FALSE,
                  leg_range = c(.7,1),
                  label_locs = NA)

ggsave("raw_data/figures/p_asymptmild_europe.pdf", plot=p_asympt_mild, dpi=600)
ggsave("raw_data/figures/p_asymptmild_europe.png", plot=p_asympt_mild, dpi=600)






```

